\documentclass[11pt,a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{bm}
\usepackage{microtype}
\usepackage{enumitem}

\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\softmax}{\mathrm{softmax}}
\newcommand{\diag}{\mathrm{diag}}

\title{PropertyScanner: Retrieval-Augmented Log-Residual Valuation with Cross-Attention}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
This note documents the valuation model and evaluation protocol used in
PropertyScanner as of the current system. The pipeline is local-first and
market-anchored: we retrieve time-safe comparables, adjust them with hedonic
and rent indices, build a robust comp baseline, and predict log-residual
quantiles via a cross-attention fusion model. The model outputs are then
blended with income signals and area intelligence, and optionally calibrated
with asymmetric conformal prediction. The goal is to produce a defensible
interval around fair value, not just a point estimate.
\end{abstract}

\section{Setup and Notation}
Each listing $i$ has an observed price $y_i > 0$, a listing type
$\ell_i \in \{\text{sale}, \text{rent}\}$, and a timestamp $\tau_i$ (monthly).
Features include tabular attributes $x_i \in \R^{d_x}$, text embeddings
$t_i \in \R^{d_t}$, and optional image embeddings $v_i \in \R^{d_v}$.
In the deployed model, $d_x = 11$, $d_t = 384$, and $d_v = 512$.
The tabular vector includes location, size, floor, sentiment, and a
price-per-sqm proxy; text includes title, description, and optionally
vision-language summaries.

For a target listing $0$, we retrieve $K$ comparables
$\mathcal{C} = \{1,\ldots,K\}$ with prices $y_j$ and features
$(x_j, t_j, v_j)$. Let $r$ denote the region (city) of the target.
The high-level pipeline is:
\[
\text{retrieve comps} \rightarrow \text{time-adjust prices} \rightarrow
\text{robust baseline} \rightarrow \text{cross-attention fusion} \rightarrow
\text{quantile outputs} \rightarrow \text{income/area blend} \rightarrow
\text{calibration and evaluation}.
\]

\section{Comparable Retrieval}
We use dense text embeddings with a LanceDB index. Let $e(\cdot)$ be the text
encoder, and define the query vector
$q = e(\text{title}_0 \oplus \text{desc}_0 \oplus \text{vlm}_0)$.
Candidates are ranked by inner product (normalized cosine similarity), then
filtered by structural and policy constraints:
\begin{align}
\text{dist}(j,0) &\leq R, \\
\text{type}(j) &= \text{type}(0), \\
1-\rho \leq \frac{\text{sqm}_j}{\text{sqm}_0} &\leq 1+\rho, \\
\tau_j &\leq \tau_0.
\end{align}
This two-stage filter enforces semantic similarity and structural coherence.
To avoid train/infer drift, the retriever metadata (encoder version and VLM
policy) is fixed across runs, and comp IDs can be persisted for parity.

\section{Time Normalization}
Comparable prices are normalized to the target month using market indices.
Let $I_{r,\ell}(\tau)$ be the monthly index for region $r$ and listing type
$\ell$ (hedonic for sale, rent index for rent). The adjusted price is
\begin{equation}
 y_j^{\ast} = y_j \cdot \frac{I_{r,\ell_0}(\tau_0)}{I_{r,\ell_j}(\tau_j)}.
\end{equation}
If indices are missing, the system falls back to the raw price rather than
extrapolating. This keeps the anchor grounded in observed data.

\section{Robust Comp Baseline}
We form a robust baseline using a MAD-filtered weighted median. Let
$\{y_j^{\ast}\}$ be the adjusted comp prices and $w_j$ be similarity or
retriever weights. Define the median $m$ and median absolute deviation
$\mathrm{MAD} = \mathrm{median}(|y_j^{\ast} - m|)$. We keep only comps with
\begin{equation}
|y_j^{\ast} - m| \leq 3\,\mathrm{MAD}.
\end{equation}
Weights are normalized, $\tilde{w}_j = w_j / \sum_k w_k$, and the baseline is
the weighted median
\begin{equation}
 b = \mathrm{wmed}\left(\{y_j^{\ast}\}, \{\tilde{w}_j\}\right).
\end{equation}
This baseline is the primary anchor for pricing and prevents outliers from
skewing the model.

\section{Multimodal Encoding and Cross-Attention}
Each modality is projected into a shared hidden dimension $d$:
\begin{align}
 h_x &= \phi_x(W_x x + b_x), \\
 h_t &= W_t t, \\
 h_v &= W_v v, \\
 h &= \phi_f([h_x; h_t; h_v]),
\end{align}
with $d = 64$ in the deployed model. For the target and its comps, define
$h_0 \in \R^d$ and $H \in \R^{K \times d}$. Cross-attention uses the target as
query and comps as keys/values:
\begin{align}
 A &= \softmax\left(\frac{h_0 H^\top}{\sqrt{d}}\right) \in \R^{1 \times K}, \\
 z &= A H \in \R^{1 \times d}, \\
 r &= g(h_0 + z),
\end{align}
where $g(\cdot)$ is a small MLP. The attention weights $A$ provide an
interpretable view of which comps drove the prediction.

\section{Log-Residual Quantile Prediction}
The model predicts log-residuals over the comp baseline. Define the target
adjusted price $y_0^{\ast}$ and residual
\begin{equation}
 \rho = \log y_0^{\ast} - \log b.
\end{equation}
For quantile $q \in \{0.1, 0.5, 0.9\}$, the model outputs $\hat{\rho}_q$ and
price quantiles are reconstructed as
\begin{equation}
 \hat{y}_q = \exp(\log b + \hat{\rho}_q).
\end{equation}
When fusion fails, the system falls back to the robust baseline with a
dispersion-based interval. This keeps valuations available while preserving a
clear provenance of the estimate.

\section{Auxiliary Heads (Rent and Time-to-Sell)}
The fusion model also emits auxiliary quantiles for rent and time-to-sell as
supporting signals. The rent and time heads output residuals $\hat{u}_q$.
Let the attention anchor be
$a = \sum_{j=1}^K A_j y_j^{\ast}$. The rent anchor is
\begin{equation}
 r_0 = 0.004\,a,
\end{equation}
with rent quantiles modeled as
\begin{equation}
 \hat{r}_q = r_0 + \sigma_r \hat{u}_q, \quad \sigma_r = 0.2 r_0.
\end{equation}
For time-to-sell, the anchor is the mean days-on-market of comps (or 90 days
when unavailable), with
\begin{equation}
 \hat{d}_q = d_0 + 30\,\hat{u}_q.
\end{equation}
In the current pipeline these heads are not trained with explicit labels, so
they are treated as optional signals rather than core valuation drivers.

\section{Training Objective}
We train on residual targets rather than raw prices. For listing $i$, the
ground-truth residual is
\[
\rho_i = \log y_i^{\ast} - \log b_i,
\]
and the model predicts residual quantiles $\hat{\rho}_{i,q}$. We use weighted
quantile regression with the pinball loss:
\begin{equation}
\ell_q(\rho_i, \hat{\rho}_{i,q}) = \max\left(q(\rho_i - \hat{\rho}_{i,q}), (q-1)(\rho_i - \hat{\rho}_{i,q})\right).
\end{equation}
The total loss with label weights $\omega_i$ is
\begin{equation}
\mathcal{L} = \frac{1}{N} \sum_{i=1}^N \omega_i \cdot
\frac{1}{3} \sum_{q \in \{0.1,0.5,0.9\}} \ell_q(\rho_i, \hat{\rho}_{i,q}).
\end{equation}
Weights reflect label reliability (sold $>$ rent ask $>$ sale ask). Training
uses AdamW with gradient clipping and early stopping. Data splits can be
standard holdout or time+geo splits to reduce leakage.

\section{Income Blend and Area Adjustment}
After fusion, we blend the model value with an income-based value and apply
area intelligence adjustments.

\paragraph{Income value.}
Let $\hat{r}$ be the monthly rent estimate from rent comps or a local yield
distribution. Let $y_m$ be the local market yield (p50). The income value is
\begin{equation}
 v_{\text{inc}} = \frac{12\,\hat{r}}{y_m/100}.
\end{equation}
We cap the adjustment to avoid large deviations. Let comp coverage be
$c = \min(1, n_{\text{rent}}/n_{\min})$ and rent uncertainty be
$u_r \geq 0$. Define a variance factor
\begin{equation}
 \psi = \frac{1}{1 + 4 u_r^2},
\end{equation}
and blend weight $\alpha = \alpha_{\max} c \psi$, with
$\alpha = 0$ if $\alpha < \alpha_{\min}$. The blended value is
\begin{equation}
 v = (1-\alpha) v_{\text{model}} + \alpha v_{\text{inc}},
\end{equation}
with uncertainty scaled by $(1 + \alpha\,\min(u_r, 0.5))$.

\paragraph{Area adjustment.}
Let area sentiment $s \in [0,1]$, development score $d \in [0,1]$, and
confidence $\kappa \in [0,1]$. We compute
\begin{equation}
 \delta = \kappa \cdot \mathrm{clip}\left((s-0.5) w_s + (d-0.5) w_d, -c_a, c_a\right),
\end{equation}
then adjust
\begin{equation}
 v' = v (1 + \delta), \quad u' = u (1 + |\delta|).
\end{equation}
This keeps the adjustment bounded and scales it by data credibility.

\section{Uncertainty and Calibration}
Raw uncertainty is the relative half-width:
\begin{equation}
 u = \frac{\hat{y}_{0.9} - \hat{y}_{0.1}}{2 \hat{y}_{0.5}}.
\end{equation}
Intervals can be calibrated with asymmetric conformal prediction. Given
residuals
\begin{align}
 e^- &= \max(0, \hat{y}_{0.1} - y), \\
 e^+ &= \max(0, y - \hat{y}_{0.9}),
\end{align}
we compute quantiles $c^- = Q_{1-\alpha}(e^-)$ and
$c^+ = Q_{1-\alpha}(e^+)$ over a rolling window (bucketed by region, property
type, and price band), and widen the interval:
\begin{equation}
 \tilde{y}_{0.1} = \hat{y}_{0.1} - c^-, \quad
 \tilde{y}_{0.9} = \hat{y}_{0.9} + c^+.
\end{equation}
Asymmetric calibration matches market intuition: downside and upside errors
behave differently in illiquid markets.

\section{Evaluation Protocol}
Evaluation targets both accuracy and risk communication. Point accuracy uses
$\hat{y}_{0.5}$:
\begin{align}
\mathrm{MAE} &= \frac{1}{N} \sum_{i=1}^N |y_i - \hat{y}_{i,0.5}|, \\
\mathrm{MAPE} &= \frac{100}{N} \sum_{i=1}^N
\frac{|y_i - \hat{y}_{i,0.5}|}{\max(y_i, 1)}, \\
\mathrm{MedAE} &= \mathrm{median}_i |y_i - \hat{y}_{i,0.5}|.
\end{align}
Interval quality is evaluated by coverage and width:
\begin{align}
\mathrm{Coverage} &= \frac{1}{N} \sum_{i=1}^N
\mathbf{1}\{\tilde{y}_{i,0.1} \le y_i \le \tilde{y}_{i,0.9}\}, \\
\mathrm{Width} &= \frac{1}{N} \sum_{i=1}^N
(\tilde{y}_{i,0.9} - \tilde{y}_{i,0.1}).
\end{align}
To reduce leakage, we emphasize time+geo splits and enforce time-safe
comparable selection. Retriever policy is frozen between training and
inference, and comp IDs can be persisted to ensure parity.

\section{Closing Perspective}
The model is intentionally modest in parameter count but explicit about market
structure. Retrieval ensures similarity, hedonic normalization keeps prices in
the correct regime, and log-residual prediction enforces anchoring to comps.
The income blend and area adjustment add domain intelligence on top of the
learned model, while conformal calibration turns predictions into intervals
that are defensible in practice. That combination is the core engineering
thesis of PropertyScanner.

\end{document}
